---
title: "p8105_hw1_sjk2254"
output: github_document
author: "Sharon Kulali"
date: "2023-09-18"
---


# Problem 1


```{r setup, message = FALSE}
# loading the needed packages

library(moderndive)
library(tidyverse)

# loading the needed data

data(early_january_weather)
```

```{r, eval = FALSE}
# assessing the data set

?early_january_weather
```

This dataset `early_january_weather` shows the hourly meteorological data for LGA, JFK, and EWR from January 2013. There are `r nrow(early_january_weather)` rows representing the hourly measurements and `r ncol(early_january_weather)` columns representing the variables which include `origin`, `year`, `month`, `day`, `hour`, `temp`, `dewp`, `humid`, `wind_dir`, `wind_speed`, `wind_gust`, `precip`, `pressure`, `visib`, and `time_hour`. Overall, the median precipitation found was `r median(pull(early_january_weather, precip))` inches with a mean temperature of `r mean(pull(early_january_weather,temp))` degrees Fahrenheit, and wind spreeds up to `r max(pull(early_january_weather, wind_speed))` mph. Notably, the wind gust variable had `r sum(is.na(pull(early_january_weather, wind_gust)))` missing values.

```{r}
# creating the scatterplot

ggplot(data = early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() + ggtitle("Temp vs Time (Hours)")
```

This scatterplot generally represents a weak, positive pattern. There appears to be a constant fluctuation from increase to decrease in temperature throughout the month but on the broader scale, you can observe that as the days progress, the temperatures continue to rise hence showing a positive relationship. Additionally, the darker shade of blue represents less humidity while the lighter shade of blue represents more humidity.

```{r}
# saving the scatterplot

ggsave("scatter_plot_p1.pdf", height = 4, width = 6)
```


# Problem #2


```{r}
# creating a data frame

p2_df = tibble(
     num_samp = rnorm(10),
     log_vec = num_samp > 0,
     chr_vec = c("female","female","unisex","unisex","male","male","female","male","unisex","female"),
     fct_vec = factor(chr_vec)
)
```

```{r}
# computing the means

m_num_samp = mean(pull(p2_df, num_samp))
m_log_vec = mean(pull(p2_df, log_vec))
m_chr_vec = mean(pull(p2_df, chr_vec))
m_fct_vec = mean(pull(p2_df, fct_vec))
```

When computing the means, it works for the sample (numeric) variable and the logical vector but not the character and factor vectors.

```{r, eval = FALSE}
# converting the variables to be numeric

an_log_vec = as.numeric(pull(p2_df,log_vec))
mean(an_log_vec)

an_chr_vec = as.numeric(pull(p2_df,chr_vec))
mean(an_chr_vec)

an_fct_vec = as.numeric(pull(p2_df,fct_vec))
mean(an_fct_vec)
```

When converting the variables to be numeric, it changes their classification types to be "numeric" instead of their previous classes of "logical", "character", or "factor". This happens because initially, R automatically classified the objects based on the inputs but now it is being clearly stated as a specific type. This explains what happens when we try to take the mean because it can only be performed on a numeric type data. More specifically, depending on how the data are stored. For example, when we change the factor vector to a numeric class, we are able to perform the mean because factors are stored as integers. On the other hand, even though we changed the character vector into a numeric class, we still cannot perform the mean because NAs are produced as a result. 